Sketch Recognition and Understanding

### 2009-08-15 draft/System.tex
after adding most cites

\subsection{Support Vector Machines}
\label{Sec:SVMdetail}
The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{} in 1995. It is gainig increase popularity due to the high performance and efficient implementation. The goal of SVM algorithms is to generate a sperating plane between two different classes which are represented using a set of n-dimentional vectors. The generated plane is used as a good classifier to the unseen exmaples. However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{}. To calcuate the margin, two different hyperplanes are constructed, one plane on each side of the data, these two planes are then moved to generenate the maximum margin possilbe.

Let us consider  a two class classification task with datapoints $x_i$ where $i=1,\cdots,m$ having corresponding labels $y_i=\pm1$ where if  $y_i=1$ if $x_i$ belonged to first class and $y_i=-1$ if $x_i$ belonged to second class. Each data  point $x_i$ is represented in $N$ dimensional input or attribule space. Let the classification plane be represented as function . 
\begin{equation}
 y=f(x)=sign(\omega \cdot x - b)
\label{eq:planeEq}
\end{equation}

 The vector $\omega$ determines th orientation of a seperation plane. The scalar $b$ determines the offeset of the plane from the orgin. Lets assume that the datatpointa re linearly seperable, i.e. there are infinitely many possible seperating planes that correctly classify the training data. Figure\ref{} illustrates two different seperating planes. It is intuitively noted from the figure that the () line is more likely to generalize more data and thus introduce less classification error. This can be geometrically characterized as plane furthest from both classes.     % Which one is preferable ? Intuitively one perfers the solid pane since small perturbations of any point would not introduce misclassifications errors. Without any information, the solid plane is more likely to generalize better on future data. Geometrically we can characterize the solid plane as being furthuret from both classes.

The SVM problem is how to construct the plane that is furthuest from both classes. One approcah is to maximize the margin between two parallel planes supporting. A plane supports a class if all points in that class are on one side of that plane. For the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i\> b$ or $\omega \cdot x_i -b \> 0$  depending on the class label. Let us suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ iss $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$. The argument insidee the decision function is invariant under positive rescalling so we will implicity fix a scale by requiring $\omega \cdot x_i -b \geq 1 $. For the points with class label $-1$ we similary require $\omega \cdot x_i -b \leq -1 $. To find the plane furthest from both setss, we can simply maximize the distance or margin between the support planes for each class as illustrated in Figure \ref{}. The support planes are pussshed apart untill they bump inot small number of data points ( the support vecotrs ) for each class. The support vectors in Figure \ref{} are outlined in (),

@@

\subsubsection{Theoretcical foundation}
From a statistical learning theory perspective these QP formulation are well founded Rougly, Statistical Learning proves that bounds on the generalization error on future points not in the training set can be obtained. Theses bounds are a function fo misclassiffied error on the training data and tersm sthat measure the complexity or capacity fo classifcation function . For linear functions, maximizing the margin of seperation as discussed abouve reduces th efunction capacilty or complexity. Thus, by explicity maximizing the margin we are minimizing bound on the generalition error and can expect better generalization with high probability. The size of the margin is not directly dependent on the dimensionaliuty of the data. Thus, we can expect good performance event for very high dimensional data (i.e, with a very large number of attribute). In a sense, problems caused by oer fitting of high - dimensional data are greatly reduced. The problem is refered to the large volume of literature in, e.g., \cite{} for more technical discussion of the taht statistical theroy.

@@

\subsubsection{Linearly Inseperable Case}
 So far we have assumed that the two datasets are linearly separable. Th this is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail. As illustrated in Figure \ref{}, if the points are not linearly seperable the primal supporitng plane will fails. Since the QP task is not feasible for ethe linearly inseperable case , then we need to relax the constrains. Consider the linear inseperable problem in Figure\ref{}, Ideally we would like no points to be missclassified and no points to fall in the margin. But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane. Any point fallihng on the wrong side of tis supprting plane is considered to be an error. We want to maximize the margin and minimize the error.

@@

Exampine what happens when athe nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define : $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize

\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_j\vartheta(x_i)\vartheta(x_j)}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
 C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
 \end{array} 
\label{eq:dualkernel}
\end{equation}
Notice that the mapped data only occurs as inner product in the objective. Now we apply a little mathematically rigorous magie known as hilbern schmidt kernels, first applied to SVMs in \cite{}.  By Mercer's theorem, we knoew that for ccertain mapping $\theta$ and any two points $u$ and $v$, the inner product of the the mapped points can be evalutated using the kernel function without ever explicity knowing ht emapping , e.g., $\theta(u)\theta(v)\equiv K(u,v)$. Some of the more popular known kernels are given below. New kernel are being developed to fit domian specific requriments. 
\begin{center}
	\begin{tabular}{cc}
	$\vartheta(u)$  & $K(u,v)$ \\ \hline 
	Degree $d$ polynomial & $\left(u\cdot v +1\right)^d$ \\ 
	Radial Basis Function & $\exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) $ \\ 
		Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\

	\end{tabular}
\end{center}

%\begin{equation}
%\begin{array}{cc}
%\underline{&}\\  
% & \left(u\cdot v +1\right)^d \\
%Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
%Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
% \end{array} 
%\label{eq:diffkernel}
%\end{equation}

@@

\subsubsection{ SVM in Machine Learning }

@@

\subsubsection {Multi Class Classifiers}
\subsubsection {SVM Implementations}
\subsubsection {Summary Of SVM}

>>>>

\subsection {Trainig the SVM}
\subsection {Testing}

\subsection{Support Vector Machines}
\label{Sec:SVMdetail}
The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{svmintroduce} in 1995. It is gainig increase popularity due to the high performance and efficient implementation. The goal of SVM algorithms is to generate a sperating plane between two different classes which are represented using a set of n-dimentional vectors. The generated plane is used as a good classifier to the unseen exmaples. However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{fig:svm1}. To calcuate the margin, two different hyperplanes are constructed, one plane on each side of the data, these two planes are then moved to generenate the maximum margin possilbe.  
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm1.jpg.eps}
	\caption{Different Classifier planes }
	\label{fig:svm1}
\end{figure}

Let us consider  a two class classification task with datapoints $x_i$ where $i=1,\cdots,m$ having corresponding labels $y_i=\pm1$ where if  $y_i=1$ if $x_i$ belonged to first class and $y_i=-1$ if $x_i$ belonged to second class. Each data  point $x_i$ is represented in $N$ dimensional input or attribule space. Let the classification plane be represented as function. 
\begin{equation}
 y=f(x)=sign(\omega \cdot x - b)
\label{eq:planeEq}
\end{equation}

 The vector $\omega$ determines th orientation of a seperation plane. The scalar $b$ determines the offeset of the plane from the orgin. Lets assume that the datatpointa re linearly seperable, i.e. there are infinitely many possible seperating planes that correctly classify the training data. Figure\ref{fig:svm1} illustrates two different seperating planes. It is intuitively noted from the figure that the solid line is more likely to generalize more data and thus introduce less classification error. This can be geometrically characterized as plane furthest from both classes.     % Which one is preferable ? Intuitively one perfers the solid pane since small perturbations of any point would not introduce misclassifications errors. Without any information, the solid plane is more likely to generalize better on future data. Geometrically we can characterize the solid plane as being furthuret from both classes.

The SVM problem is how to construct the plane that is furthuest from both classes. One approcah is to maximize the margin between two parallel planes supporting. A plane supports a class if all points in that class are on one side of that plane. For the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i\> b$ or $\omega \cdot x_i -b \> 0$  depending on the class label. Let us suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ iss $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$. The argument inside the decision function is invariant under positive rescalling so we will implicity fix a scale by requiring $\omega \cdot x_i -b \geq 1 $. For the points with class label $-1$ we similary require $\omega \cdot x_i -b \leq -1 $. To find the plane furthest from both setss, we can simply maximize the distance or margin between the support planes for each class as illustrated in Figure \ref{fig:svm2}. The support planes are pussshed apart untill they bump inot small number of data points ( the support vecotrs ) for each class. The support vectors in Figure \ref{fig:svm2} are filled datapoints , 
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm2.jpg.eps}
	\caption{Margin of Support planes }
	\label{fig:svm2}
\end{figure}

@@

\subsubsection{Theoretcical foundation}
From a statistical learning theory perspective these QP formulation are well founded Rougly, Statistical Learning proves that bounds on the generalization error on future points not in the training set can be obtained. Theses bounds are a function fo misclassiffied error on the training data and tersm sthat measure the complexity or capacity fo classifcation function . For linear functions, maximizing the margin of seperation as discussed abouve reduces th efunction capacilty or complexity. Thus, by explicity maximizing the margin we are minimizing bound on the generalition error and can expect better generalization with high probability. The size of the margin is not directly dependent on the dimensionaliuty of the data. Thus, we can expect good performance event for very high dimensional data (i.e, with a very large number of attribute). In a sense, problems caused by oer fitting of high - dimensional data are greatly reduced. The problem is refered to the large volume of literature in, e.g., \cite{bookSVMoverfit11,statisticalSVM1} for more technical discussion of the taht statistical theroy.

@@

\subsubsection{Linearly Inseperable Case}
 So far we have assumed that the two datasets are linearly separable. This is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail. As illustrated in Figure \ref{fig:svm5}, if the points are not linearly seperable the primal supporitng plane will fails. Since the QP task is not feasible for ethe linearly inseperable case , then we need to relax the constrains. Consider the linear inseperable problem in Figure\ref{fig:svm5}, Ideally we would like no points to be missclassified and no points to fall in the margin. But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane. Any point fallihng on the wrong side of tis supprting plane is considered to be an error. We want to maximize the margin and minimize the error.

\begin{figure}
	\centering
		\includegraphics{afterDefense/svm5.jpg.eps}
	\caption{Linearly Inseperable}
	\label{fig:svm5}
\end{figure}

@@

Examine what happens when athe nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define : $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize

\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_j\vartheta(x_i)\vartheta(x_j)}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
 C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
 \end{array} 
\label{eq:dualkernel}
\end{equation}
Notice that the mapped data only occurs as inner product in the objective. Now we apply a little mathematically rigorous magie known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.  By Mercer's theorem, we knoew that for ccertain mapping $\theta$ and any two points $u$ and $v$, the inner product of the the mapped points can be evalutated using the kernel function without ever explicity knowing ht emapping , e.g., $\theta(u)\theta(v)\equiv K(u,v)$. Some of the more popular known kernels are given below. New kernel are being developed to fit domian specific requriments. 
\begin{center}
	\begin{tabular}{cc}
	$\vartheta(u)$  & $K(u,v)$ \\ \hline 
	Degree $d$ polynomial & $\left(u\cdot v +1\right)^d$ \\ 
	Radial Basis Function & $\exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) $ \\ 
		Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\

	\end{tabular}
\end{center}
To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective insted of the original dot product. Thus by changing kernels we can get different hightly nonlinear classifiers. No algorithmic changes are required from linear clase other than substitue of a kernel evaluation for simple dot product. All the benefits fo the original linear SVM methods are maintained. We can train a highly nonlinear classification function such as a ploynomial or a raadial basis function machine, or a sigmoidal neural network using robust, efficient algorithsm that have no problem with local minima. By using kernel substitution a linear algorithm (only capable of handling seperable data) can be turned into a general nonlinear algorithm. 
%\begin{equation}
%\begin{array}{cc}
%\underline{&}\\  
% & \left(u\cdot v +1\right)^d \\
%Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
%Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
% \end{array} 
%\label{eq:diffkernel}
%\end{equation}

@@

\subsubsection{ SVM in Machine Learning }
SVM has been successfully applied to a number of applications ranging from particle identification \cite{ParticleSVM7}. Face detection  \cite{faceSVM8,faceSVM7} and text categorization \cite{TextSVM7} to engine knock detection. SVMs are very suited to the classification task that commonly arise in machine vision. As an example we consider and application involving face identification\cite{faceSVM8}.  Three different SVM classiifer were used on training set of 200 images (5 image per person). The error rate was less than 5\%. An RBF network and SVM classifiers were compared on the UPUS handwritten digit recognition database\cite{ORLDataset}.  SVM outperformed and RBF network on all digits.

@@

SVM have been applied ot the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{IjdarSherifPaper}. Sherif and Ezzat has performed a comparision between SVM and several other techniques on the MNIST dataset. They also compared the same classfication methods on arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}. Their results shows that SVM classifier outperform all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%. Similar results was reported by \cite{empiricalcomp11,SVMInvariantComDecoste02} on the MNIST dataset. Ther are many other successful application SVM. They have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Schölkopf97comparingsupport}.

This seperior performance because they may elimnates many of problem experienced with other inference methodoliges with other inference methodollogies like neural network and decision trees. The first problem that is eliminated was the problems with local minima. SVM can construct highly nonlinear classification and regression function without woring about getting stuck at local minima. Another problem SVM eliminated is the number of parameter that are picked ot generate the model. For example, SVM classifier with RBF  kernel will only need 2 paramters to set the model the $C$ cost paramter and $\sigma$ paramter of the RBF kernel. The last reason is the final result are sable, reproducible and largely independent of the specific algorithm used to  optimize the SVM model. If two users apply the same SVM model with the same parameters to the same data, they will et the same solution modeule numeric issues. Compare this with neural networks where the results are dependent on the particular algorithms and starting point used.

SVM classifiers are also simple to use. One need not be a SVM expert to successfully apply existing SVM software to new problems.

%different classifier on MNIST and a

\subsubsection {Multi Class Classifiers}

\subsubsection {SVM Implementations}
Typically an SVM approach requires the solution of a QP or LP problem. LP and QP type problems have been extensively studied inthe field of mathematical programming. These can be divided into three categories techniques in which kernel compoinents are evaluated and discarded during learning. Decomposition methods in which an evolving subset of data is used and new optimization approaches that specifically exploit the structure of the SVM problem.

For the first category the most obvioius  approach is to sequentially update the $\alpha_i$  and this is the approach used by the kernel  Adtron (KA) algorithm \cite{}.

Chunking and decomposition methods optimize the SVM with respect to subsets. Rather than sequentially updateing the $\alpha_i$ the alternative is to upodate the $\alpha_i$ in parallel but using only a subset or wokring set of data at each stage. There are many successful code based on these decomposition strategies. SVM codes available online such as SVMTorch \cite{} and SVMLight \cite{} use these working set strategies.

The third approach is to directly attack the SVM  problem from and optimization prespective and create algorithms that explicity exploits the structure of teh problem. Keerthi et al. \cite{} proposed a very effective algorithm based on the dual geometry of finding the two closest points in the convex hulls.

\subsubsection {Summary Of SVM}
